---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/graphicalMCP)](https://cran.r-project.org/package=graphicalMCP)
[![R-CMD-check](https://github.com/Gilead-BioStats/graphicalMCP/workflows/R-CMD-check-main/badge.svg)](https://github.com/Gilead-BioStats/graphicalMCP/actions)
[![Codecov test coverage](https://codecov.io/gh/Gilead-BioStats/graphicalMCP/branch/s3-graph_mcp/graph/badge.svg)](https://app.codecov.io/gh/Gilead-BioStats/graphicalMCP?branch=s3-graph_mcp)
[![R-CMD-check](https://github.com/Gilead-BioStats/graphicalMCP/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Gilead-BioStats/graphicalMCP/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Introduction

A multiple comparison procedure (MCP) is a statistical analysis method that allows for assessing the efficacy of multiple endpoints, some of which are dependent on each other, in a single clinical trial. Endpoints can be different doses, treatment of different conditions, or combined superiority & non-inferiority testing.

In [Bretz et al (2011)](https://onlinelibrary.wiley.com/doi/10.1002/bimj.201000239), a graphical method to MCPs was described, which separates the weighting of the dependency graph from the particular statistical test used to assess each endpoint. This package is a low-dependency implementation of those methods.

## Installation

graphicalMCP is not on CRAN, so install it from GitHub with

```{r eval=FALSE}
# install.packages("pak")
pak::pak("Gilead-BioStats/graphicalMCP")
```

## Basic usage

### Initial graph

The base object in graphicalMCP is an `initial_graph`, which is a weighted, directed graph represented by a matrix of transition (edge) weights, and a vector of hypothesis (vertex) weights.

```{r create-graph}
library(graphicalMCP)

# A graphical multiple comparison procedure with two primary hypotheses (H1
# and H2) and two secondary hypotheses (H3 and H4)
# See Figure 1 in Bretz, F., Posch, M., Glimm, E., Klinglmueller, F., Maurer,
# W., & Rohmeyer, K. (2011). Graphical approaches for multiple comparison
# procedures using weighted Bonferroni, Simes, or parametric tests. Biometrical
# Journal, 53(6), 894-913.
hypotheses <- c(0.5, 0.5, 0, 0)
transitions <- rbind(
  c(0, 0, 1, 0),
  c(0, 0, 0, 1),
  c(0, 1, 0, 0),
  c(1, 0, 0, 0)
)
names <- c("A1", "A2", "B1", "B2")
g_dose <- create_graph(hypotheses, transitions, names)

g_dose
```

### Update graph

Hypotheses can be rejected from the MCP using `update_graph()`. Updated weights and transitions are calculated according to the weighting strategy in [Bretz et al (2011)](https://onlinelibrary.wiley.com/doi/10.1002/bimj.201000239)

```{r update-graph}
update_graph(g_dose, c(TRUE, FALSE, FALSE, TRUE))
```

### Generate weights

The weights of all sub-graphs can be calculated with `generate_weights()`. This uses more efficient code under the hood than `update_graph()` in order to be performant for larger graphs.

```{r generate-weights}
generate_weights(g_dose)
```

### Test hypotheses

A mixture of statistical tests are (or will be) supported in graphicalMCP. A graph can be tested against a given alpha with `test_all_subgraphs()`. A report is then generated, showing the graph & test results.

In this example, a weighted Bonferroni test is applied to all hypotheses, with a threshold of 0.05. We can reject a given intersection hypothesis if any of the individual hypotheses within that sub-graph passes the test assigned to it. We can then reject a hypothesis globally if all intersection hypotheses containing that hypothesis are rejected. For instance, in this example we can reject every intersection hypothesis except the one containing only B1 & B2 (Row 3 of `test_results`). Thus, we can reject the null hypotheses for A1 & A2, but we cannot reject the null hypotheses for B1 & B2.

```{r test-graph}
test_all_subgraphs(
  g_dose,
  p_values = c(.01, .02, .03, .05),
  alpha = .05,
  tests = list(
    bonferroni = list(1:4),
    simes = list(),
    parametric = list()
  )
)
```

## Related work

These methods were originally implemented in R after the 2011 paper in the gMCP package, which is still available on CRAN today:

- `install.packages("gMCP")`
- https://github.com/kornl/gMCP

However, because development has ceased on this original package, we hope to re-implement the methods with a clearer distinction between weighting procedures and test procedures; with fewer dependencies, in particular shedding the Java dependency; with the simpler, more transparent S3 class framework; and with improvements to the accuracy of the parametric tests method.
