---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/graphicalMCP)](https://cran.r-project.org/package=graphicalMCP)
[![R-CMD-check](https://github.com/Gilead-BioStats/graphicalMCP/workflows/R-CMD-check-main/badge.svg)](https://github.com/Gilead-BioStats/graphicalMCP/actions)
[![Codecov test coverage](https://codecov.io/gh/Gilead-BioStats/graphicalMCP/branch/s3-graph_mcp/graph/badge.svg)](https://app.codecov.io/gh/Gilead-BioStats/graphicalMCP?branch=s3-graph_mcp)
[![R-CMD-check](https://github.com/Gilead-BioStats/graphicalMCP/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Gilead-BioStats/graphicalMCP/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Introduction

A multiple comparison procedure (MCP) is a statistical analysis method that allows for assessing the efficacy of multiple endpoints, some of which are dependent on each other, in a single clinical trial. Endpoints can be different doses, treatment of different conditions, or combined superiority & non-inferiority testing.

In [Bretz et al (2011)](https://onlinelibrary.wiley.com/doi/10.1002/bimj.201000239), a graphical method to MCPs was described, which separates the weighting of the dependency graph from the particular statistical test used to assess each endpoint. This package is a low-dependency implementation of those methods.

## Installation

graphicalMCP is not on CRAN, so install it from GitHub with

```{r eval=FALSE}
# install.packages("pak")
pak::pak("Gilead-BioStats/graphicalMCP")
```

## Basic usage

The base object in graphicalMCP is an `mcp_graph`, which is a weighted, directed graph represented by a matrix of transition (edge) weights, and a vector of hypothesis (vertex) weights.

```{r}
library(graphicalMCP)

transitions <- rbind(
  c(0, .5, .5),
  c(.5, 0, .5),
  c(.5, .5, 0)
)
hypotheses <- rep(.333333, 3)

g_dose <- create_graph(hypotheses, transitions, paste("dose", letters[1:3]))

g_dose
```

## Related work

These methods were originally implemented in R after the 2011 paper in the gMCP package, which is still available on CRAN today:

- `install.packages("gMCP")`
- https://github.com/kornl/gMCP

However, because development has ceased on this original package, we hope to re-implement the methods with a clearer distinction between weighting procedures and test procedures; with fewer dependencies, in particular shedding the Java dependency; with the simpler, more transparent S3 class framework; and with improvements to the accuracy of the parametric tests method.
