---
title: "Common multiple comparison procedures illustrated using graphicalMCP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Common procedures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.align = "center",
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(5, 5)
)
```

```{r setup}
library(graphicalMCP)
```

## Introduction

In confirmatory clinical trials, regulatory guidelines mandate the strong control 
of the family-wise error rate at a prespecified level $\alpha$. Many multiple comparison 
procedures (MCPs) have been proposed for this purpose. The graphical approaches 
are a general framework that include many common MCPs as special cases. In this 
vignette, we illustrate how to use graphicalMCP to perform some common MCPs.

In graphical approaches to test $m$ hypotheses, each hypothesis $H_i$ receives a 
weight $0\leq w_i\leq 1$ (called hypothesis weight), where $\sum_{i=1}^{m}w_i\leq 1$. 
From $H_i$ to $H_j$, there could be a directed and weighted edge $0\leq g_{ij}\leq 1$, 
which means that when $H_i$ is rejected, its hypothesis weight will be propagated 
(or transferred) to $H_j$ and $g_{ij}$ determines how much of the propagation. We 
also require $\sum_{j=1}^{m}g_{ij}\leq 1$ and $g_{ii}=0$.

## Bonferroni-based procedures
### Weighted Bonferroni test

A weighted Bonferroni test splits $\alpha$ among hypotheses by testing every hypothesis 
at a significance level of $w_i\alpha$. Thus it rejects a hypothesis if its p-value 
is less than or equal to its significance level. When $w_i=w_j$ for all $i,j$, this 
means an equal split and the test is the Bonferroni test. There is no propagation 
between any pair of hypothesis.

### error in plotting
### Error in round(edge_labels, precision) : 
###  non-numeric argument to mathematical function
```{r bonferroni}
set.seed(1234)
alpha <- 0.025
m <- 3
transitions <- matrix(0, m, m)
bonferroni_graph <- graph_create(rep(1 / m, m), transitions)
plot(bonferroni_graph, layout = igraph::layout_in_circle, vertex.size = 30)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(bonferroni_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

### Holm Procedure

Holm (or Bonferroni-Holm) procedures improve over Bonferroni tests by allowing 
propagation. In other words, transition weights between hypotheses may not be zero. 
So it is uniformly more powerful than Bonferroni tests.

```{r bonferroni-holm}
set.seed(1234)
alpha <- 0.025
m <- 3
transitions <- matrix(1 / (m - 1), m, m)
diag(transitions) <- 0
holm_graph <- graph_create(rep(1 / m, m), transitions)
plot(holm_graph, layout = igraph::layout_in_circle, vertex.size = 30)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(holm_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

### Fixed sequence procedure

Fixed sequence (or hierarchical) procedures pre-specify an order of testing. For 
example, the procedure will test $H_1$ first. If it is rejected, it will test $H_2$; 
otherwise the testing stops. If $H_2$ is rejected, it will test $H_3$; otherwise 
the testing stops. For each hypothesis, it will be tested at the full $\alpha$ level, 
when it can be tested.

```{r fixed-sequence, fig.dim=c(6, 2)}
set.seed(1234)
alpha <- 0.025
m <- 3
transitions <- rbind(
  c(0, 1, 0),
  c(0, 0, 1),
  c(0, 0, 0)
)
fixed_sequence_graph <- graph_create(c(1, 0, 0), transitions)
plot(fixed_sequence_graph, nrow = 1, asp = .1)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(fixed_sequence_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

### Fallback procedure

Fallback procedures have one-way propagation (like fixed sequence procedures) but 
allow hypotheses to be tested at different significance levels.

```{r fixed-sequence, fig.dim=c(6, 2)}
set.seed(1234)
alpha <- 0.025
m <- 3
# Wiens (2003)
transitions <- rbind(
  c(0, 1, 0),
  c(0, 0, 1),
  c(0, 0, 0)
)
fallback_graph_2003 <- graph_create(rep(1 / m, m), transitions)
plot(fallback_graph_2003, nrow = 1, asp = .1)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(fallback_graph_2003, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

Further they can be improved to allow propagation from later hypotheses to earlier 
hypotheses, because it is possible that a later hypothesis is rejected before an earlier 
hypothesis can be rejected.

```{r fixed-sequence-improved, fig.dim=c(6, 2)}
set.seed(1234)
alpha <- 0.025
m <- 3
# Wiens and Dmitrienko (2005)
transitions <- rbind(
  c(0, 1, 0),
  c(0, 0, 1),
  c(0.5, 0.5, 0)
)
fallback_graph_2005 <- graph_create(rep(1 / m, m), transitions)
plot(fallback_graph_2008, nrow = 1, asp = .1)

# Hommel and Bretz (2008)
epsilon <- 0.0001
transitions <- rbind(
  c(0, 1, 0),
  c(1 - epsilon, 0, epsilon),
  c(1, 0, 0)
)
fallback_graph_2008 <- graph_create(rep(1 / m, m), transitions)
plot(fallback_graph_2008, nrow = 1, asp = .1)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(fallback_graph_2008, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

### Serial gatekeeping procedure

Serial gatekeeping procedures involve ordered multiple families of hypotheses, 
where all hypotheses of a family of hypotheses must be rejected before proceeding 
in the test sequence. The example below considers a primary family consisting of two 
hypotheses $H_1$ and $H_2$ and a secondary family consisting of a single hypothesis 
$H_3$. In the primary family, the Holm procedure is applied. If both $H_1$ and $H_2$ 
are rejected, $H_3$ can be tested at level $\alpha$; otherwise $H_3$ cannot be rejected. 
To allow the conditional propagation to $H_3$, an $\varepsilon$ edge is used from 
$H_2$ to $H_3$. It has a very small transition weight so that $H_2$ propagates most
of its hypothesis weight to $H_1$ (if not already rejected) and retains a small 
(non-zero) weight for $H_3$ so that if $H_1$ has been rejected, all hypothesis 
weight of $H_2$ will be propagated to $H_3$. Here $\varepsilon$ is assigned to be 
0.0001 and in practice, the value could be adjusted but it should be much smaller 
than the smallest p-value observed.

```{r serial-gatekeeping, fig.dim=c(6, 2)}
set.seed(1234)
alpha <- 0.025
m <- 3
epsilon <- 0.0001
transitions <- rbind(
  c(0, 1, 0),
  c(1 - epsilon, 0, epsilon),
  c(0, 0, 0)
)
serial_gatekeeping_graph <- graph_create(c(0.5, 0.5, 0), transitions)
plot(serial_gatekeeping_graph, nrow = 1, asp = .1)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(serial_gatekeeping_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

### Parallel gatekeeping procedure

Parallel gatekeeping procedures also involves multiple ordered families of hypotheses, 
where any null hypotheses of a family of hypotheses must be rejected before proceeding 
in the test sequence. The example below considers a primary family consisting of two 
hypotheses $H_1$ and $H_2$ and a secondary family consisting of two hypotheses $H_3$ 
and $H_4$. In the primary family, the Bonferroni test is applied. If any of $H_1$ 
and $H_2$ is rejected, $H_3$ and $H_4$ can be tested at level $\alpha/2$ using the 
Holm procedure; if both $H_1$ and $H_2$ are rejected, $H_3$ and $H_4$ can be tested 
at level $\alpha$ using the Holm procedure; otherwise $H_3$ and $H_4$ cannot be 
rejected.

```{r parallel-gatekeeping, fig.dim=c(6, 2)}
set.seed(1234)
alpha <- 0.025
m <- 4
transitions <- rbind(
  c(0, 0, 0.5, 0.5),
  c(0, 0, 0.5, 0.5),
  c(0, 0, 0,   1),
  c(0, 0, 1,   0)
)
parallel_gatekeeping_graph <- graph_create(c(0.5, 0.5, 0, 0), transitions)
plot(parallel_gatekeeping_graph, nrow = 1, asp = .1)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(parallel_gatekeeping_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

The above parallel gatekeeping procedure can be improved by adding $\varepsilon$ 
edges from secondary hypotheses to primary hypotheses, because it is possible that 
both secondary hypotheses are rejected but there is still a remaining primary hypothesis 
not rejected.

```{r parallel-gatekeeping-improved, fig.dim=c(6, 2)}
set.seed(1234)
alpha <- 0.025
m <- 4
epsilon <- 0.0001
transitions <- rbind(
  c(0, 0, 0.5, 0.5),
  c(0, 0, 0.5, 0.5),
  c(epsilon, 0, 0,   1 - epsilon),
  c(0, epsilon, 1 - epsilon,   0)
)
parallel_gatekeeping_improved_graph <- graph_create(c(0.5, 0.5, 0, 0), transitions)
plot(parallel_gatekeeping_improved_graph, nrow = 1, asp = .1)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(parallel_gatekeeping_improved_graph, 
                                    p = p_values, alpha = alpha)
test_results$outputs$rejected
```

### Successive procedure

Successive procedures incorporate successive relationships between hypotheses. For 
example, the secondary hypothesis is not tested until the primary hypothesis has 
been rejected. This is similar to using the fixed sequence procedure as a component 
of a graph. The example below considers two primary hypotheses $H_1$ and $H_2$ and 
two secondary hypotheses $H_3$ and $H_4$. Primary hypotheses $H_1$ and $H_2$ receive 
the equal hypothesis weight of 0.5; secondary hypotheses $H_3$ and $H_4$ receive 
the hypothesis weight of 0. A secondary hypothesis $H_3 (H_4)$ can be tested only if 
the corresponding primary hypothesis $H_1 (H_2)$ has been rejected. This represents 
the successive relationships between $H_1$ and $H_3$, and $H_2$ and $H_4$, respectively. 
If both $H_1$ and $H_3$ are rejected, their hypothesis weights are propagated to 
$H_2$ and $H_4$, and vice versa.

```{r simple-successive}
set.seed(1234)
alpha <- 0.025
m <- 4
transitions <- rbind(
  c(0, 0, 1, 0),
  c(0, 0, 0, 1),
  c(0, 1, 0, 0),
  c(1, 0, 0, 0)
)
#  Maurer et al. (2011)
simple_successive_graph <- graph_create(c(0.5, 0.5, 0, 0), transitions)
plot(simple_successive_graph, layout = "grid", nrow = 2, vertex.size = 30)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(simple_successive_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

The above graph could be generalized to allow propagation between primary hypotheses. 
A general successive graph is illustrate below with a variable to determine the 
propagation between $H_1$ and $H_2$.

```{r simple-successive-var}
set.seed(1234)
alpha <- 0.025
m <- 4
#  Maurer et al. (2011)
successive_var <- simple_successive_var <- function(gamma) {
  graph_create(
    c(.5, .5, 0, 0),
    rbind(
      c(0, gamma, 1 - gamma, 0),
      c(gamma, 0, 0, 1 - gamma),
      c(0, 1, 0, 0),
      c(1, 0, 0, 0)
    )
  )
}
successive_var_graph <- successive_var(0.5)
plot(successive_var_graph, layout = "grid", nrow = 2, vertex.size = 30)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_shortcut(successive_var_graph, p = p_values, alpha = alpha)
test_results$outputs$rejected
```

## Simes-based procedures
### Hommel procedure

Hommel (1988) procedure is a closed test procedure which uses Simes tests for every 
intersection hypothesis. According to Xi and Bretz (2019), the graph for Hommel 
procedures is the same as the graph for Holm procedures. Thus to perform Hommel 
procedure, we just need to specify `test_type` to be `simes`.

```{r hommel}
set.seed(1234)
alpha <- 0.025
m <- 3
transitions <- matrix(1 / (m - 1), m, m)
diag(transitions) <- 0
hommel_graph <- graph_create(rep(1 / m, m), transitions)
plot(hommel_graph, layout = igraph::layout_in_circle, vertex.size = 30)
p_values <- runif(m, 0, alpha)
test_results <- graph_test_closure(hommel_graph, p = p_values, alpha = alpha, 
                                   test_types = "simes")
test_results$outputs$rejected
```

## parametric procedures
### Dunnett procedure

Step-down Dunnett procedures are a closed test procedure and an improvement from 
Holm procedures by incorporating the correlation structure between test statistics.
Thus they are the same graph as Holm procedures. Assume an equi-correlated case, 
where the correlation bewteen any pair of test statistics is the same, e.g., 0.5. 
Then we can perform the step-down Dunnett procedure by specifying `test_type` to 
be `parametric` and providing the correlation matrix.

```{r dunnett}
set.seed(1234)
alpha <- 0.025
m <- 3
transitions <- matrix(1 / (m - 1), m, m)
diag(transitions) <- 0
dunnett_graph <- graph_create(rep(1 / m, m), transitions)
plot(dunnett_graph, layout = igraph::layout_in_circle, vertex.size = 30)
p_values <- runif(m, 0, alpha)
corr <- matrix(0.5, m, m)
diag(corr) <- 1
test_results <- graph_test_closure(dunnett_graph, p = p_values, alpha = alpha, 
                                   test_types = "parametric",
                                   test_corr = list(corr))
test_results$outputs$rejected
```

################################################################################

<!-- This paragraph needs work -->

These procedures are popular because they control the FWER while being simple, allowing them to be easily understood and communicated about. However, they're not always the procedures which fit the study best, and using more complex procedures can make communication difficult. The graphical approach can enable more powerful procedures to be used while enabling clinical teams to understand and communicate the procedures.

# Complex procedures

## Simple successive

The simple successive procedure was introduced by [names] in their [year] paper, [*title*]. It is designed for experiments involving four hypotheses: Two pairs of hypotheses, each containing a primary and a related secondary hypothesis. Initial weight is split equally between pairs and is passed from one pair to the other only when both hypotheses in a pair are rejected.

```{r simple-successive}
simple_successive_graph <- graph_create(
  c(.5, .5, 0, 0),
  rbind(
    c(0, 0, 1, 0),
    c(0, 0, 0, 1),
    c(0, 1, 0, 0),
    c(1, 0, 0, 0)
  )
)

plot(simple_successive_graph, layout = "grid", nrow = 2, vertex.size = 30)
```

General successive graphs are a good example where multiple variations may be useful with slight differences in starting edge weights. Variable edge weights are not currently supported, but they can be done for a particular graph with a light wrapper.

```{r simple-successive-var-fun}
simple_successive_var <- function(gamma) {
  graph_create(
    c(.5, .5, 0, 0),
    rbind(
      c(0, gamma, 1 - gamma, 0),
      c(gamma, 0, 0, 1 - gamma),
      c(0, 1, 0, 0),
      c(1, 0, 0, 0)
    )
  )
}
```

Then multiple variations can be created and compared easily. A similar approach could be used for deviating from the equal split of hypothesis weights as well.

```{r simple-successive-var, fig.show="hold", fig.align='default', out.width="47%"}
plot(
  simple_successive_var(.75),
  layout = "grid",
  nrow = 2,
  vertex.label.cex = 1.4,
  edge.label.cex = 1.4,
  vertex.size = 30
)

plot(
  simple_successive_var(.9),
  layout = "grid",
  nrow = 2,
  vertex.label.cex = 1.4,
  edge.label.cex = 1.4,
  vertex.size = 30
)
```

## Huque-Alosh-Bhore (2011)

In 2011, Huque, Alosh, and Bhore proposed another size 4 procedure. What's this one particularly meant for?

```{r huque-alosh-bhore}
hab_2011_graph <- graph_create(
  c(1, 0, 0, 0),
  rbind(
    c(0, .5, .5, 0),
    c(0, 0, 0, 1),
    c(0, .5, 0, .5),
    c(0, 1, 0, 0)
  )
)

plot(hab_2011_graph, layout = "grid", vertex.size = 30)
```

## Wiens-Dmitrienko (2005)

```{r wiens-dmitrienko, fig.dim=c(6, 2)}
wd_2005_graph <- graph_create(
  c(1 / 3, 1 / 3, 1 / 3),
  rbind(
    c(0, 1, 0),
    c(0, 0, 1),
    c(.5, .5, 0)
  )
)

plot(
  wd_2005_graph,
  layout = "grid",
  nrow = 1,
  edge_curves = c(pairs = -4, "H3|H1" = -4),
  vertex.size = 30,
  asp = .1
)
```

## Parallel gate-keeping

In [year], [names] introduced the parallel gate-keeping procedure. It is useful when...?

```{r parallel-gatekeeping}
par_gate_graph <- graph_create(
  c(.5, .5, 0, 0),
  rbind(
    c(0, 0, .5, .5),
    c(0, 0, .5, .5),
    c(0, 0, 0, 1),
    c(0, 0, 1, 0)
  )
)


plot(par_gate_graph, layout = "grid", nrow = 2, vertex.size = 30)
```

## Improved parallel gate-keeping

One shortcoming of the parallel gate-keeping approach is in the case where e.g. H2, H3, and H4 are found to be significant. Once these three have been deleted, H1 is left with only 0.5 weight. This makes the graph sub-optimal.

```{r parallel-gatekeeping-update}
plot(graph_update(par_gate_graph, 2:4), vertex.size = 30)
```

The graph can be made optimal by adding epsilon edges going back up from the secondary to the primary hypotheses. Epsilon edges have some infinitesimal weight which creates a connection between hypotheses that effectively only propagates weight when it's the only edge left exiting a hypothesis. This yields the improved parallel gate-keeping procedure.

```{r improved-parallel-gatekeeping}
imp_par_gate_graph <- graph_create(
  c(.5, .5, 0, 0),
  rbind(
    c(0, 0, .5, .5),
    c(0, 0, .5, .5),
    c(.0001, 0, 0, .9999),
    c(0, .0001, .9999, 0)
  )
)

plot(
  imp_par_gate_graph,
  layout = "grid",
  nrow = 2,
  eps = .001,
  vertex.size = 30
)
```

```{r improved-parallel-gatekeeping-update, fig.show="hold", fig.align='default', out.width="30%"}
plot(
  graph_update(imp_par_gate_graph, c(2)),
  vertex.size = 60,
  vertex.label.cex = 2.5,
  edge.label.cex = 2.5,
  eps = 1e-3
)

plot(
  graph_update(imp_par_gate_graph, c(2, 4)),
  vertex.size = 60,
  vertex.label.cex = 2.5,
  edge.label.cex = 2.5
)

plot(
  graph_update(imp_par_gate_graph, c(2, 4, 3)),
  vertex.size = 60,
  vertex.label.cex = 2.5,
  edge.label.cex = 2.5
)
```

# Conclusion

These examples are just a few of the unlimited possibilities enabled by the graphical approach to MCPs. Graphs can be customized based on priorities in trial design, marginal power of the endpoints, and other assumptions, such as expected or known correlation between endpoints. The graphical approach allows these complex procedures to maximize the chances of having a successful trial while enabling better communication and insight about the procedures.
