---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  results = "hide",
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

## Introduction

A multiple comparison procedure (MCP) is a statistical analysis method that allows for assessing the efficacy of multiple endpoints, some of which are dependent on each other, in a single clinical trial. Endpoints can be different doses, treatment of different conditions, combined superiority & non-inferiority testing, or many others. A key feature is that MCPs strongly control the type-I error rate at level alpha.

In [Bretz et al (2011)](https://onlinelibrary.wiley.com/doi/10.1002/bimj.201000239), a graphical method for MCPs is described, which separates the weighting of the clinical endpoints from the particular statistical test used to assess each endpoint. A graphical approach can also be helpful for communicating study design to clinical teams. This package is a low-dependency implementation of those methods.

## Installation

graphicalMCP is not on CRAN, so install it from GitHub with

```{r eval=FALSE}
# install.packages("pak")
pak::pak("Gilead-BioStats/graphicalMCP@dev")
```

## Basic usage

### Initial graph

The base object in graphicalMCP is an `initial_graph`, which is a weighted, directed graph represented by a matrix of transition (edge) weights, and a vector of hypothesis (vertex) weights.

```{r create-graph, fig.dim=c(3, 3)}
library(graphicalMCP)

# A graphical multiple comparison procedure with two primary hypotheses (H1
# and H2) and two secondary hypotheses (H3 and H4)
# See Figure 1 in Bretz, F., Posch, M., Glimm, E., Klinglmueller, F., Maurer,
# W., & Rohmeyer, K. (2011). Graphical approaches for multiple comparison
# procedures using weighted Bonferroni, Simes, or parametric tests. Biometrical
# Journal, 53(6), 894-913.
hypotheses <- c(0.5, 0.5, 0, 0)
transitions <- rbind(
  c(0, 0, 1, 0),
  c(0, 0, 0, 1),
  c(0, 1, 0, 0),
  c(1, 0, 0, 0)
)

hyp_names <- c("H1", "H2", "H3", "H4")
example_graph <- graph_create(hypotheses, transitions, hyp_names)

plot(example_graph, layout = "grid", vertex.size = 60)
```

### Update graph

Hypotheses can be deleted from the MCP using `graph_update()`. Updated weights and transitions are calculated according to the weighting strategy in Algorithm 1 of [Bretz et al (2011)](https://onlinelibrary.wiley.com/doi/10.1002/bimj.201000239). We distinguish *deleting* from *rejecting* in the [glossary](#glossary-of-terms).

```{r update-graph}
graph_update(example_graph, delete = c(FALSE, TRUE, TRUE, FALSE))
```

### Generate weights

The weights of all sub-graphs can be calculated with `graph_generate_weights()`. This uses more efficient code under the hood than `graph_update()` in order to be performant for larger graphs.

```{r generate-weights}
graph_generate_weights(example_graph)
```

More information on the closure can be found in [Link to closure vignette].

### Test hypotheses

Bonferroni testing via the shortcut method is supported in graphicalMCP. Such a test can be performed with `graph_test_shortcut()`, which generates a report showing the graph & test results. See more insight about why a hypothesis was rejected or not by setting the `verbose` and `test_values` flags. More details about shortcut testing can be found in `vignette("shortcut-testing")`.

```{r test-graph-shortcut}
graph_test_shortcut(example_graph, p = c(.01, .03, .02, .01), alpha = .025)
```

A graph can also be tested using Simes- or parametric-based testing using the closure test. Other types of tests will be added over time, and a combination of tests can be used for groups of hypotheses. Additional details about closure testing can be found in `vignette("closed-testing")`.

```{r test-graph}
graph_test_closure(
  example_graph,
  p = c(.01, .03, .02, .01),
  alpha = .025,
  test_types = "bonferroni",
  test_groups = list(1:4)
)
```

### Power simulations

It's not always obvious from a graph structure how easy or difficult it will be to reject each hypothesis. One way to understand this better is to run a power simulation, where many p-values are simulated, and a graph is tested against each.

```{r power}
graph_calculate_power(
  example_graph,
  sim_n = 1e5,
  power_marginal = c(.9, .9, .8, .8)
)
```

All valid test types & hypothesis groupings are valid for power simulations as well. Power simulations are discussed further in both the `vignette("shortcut-testing")` and the `vignette("closed-testing")`.

## Related work

These methods were originally implemented in the [gMCP package](https://github.com/kornl/gMCP), which is still available on CRAN today. There is also a lighter version of gMCP implemented in [gMCPmini](https://github.com/allenzhuaz/gMCPmini) and its successor, [gMCPLite](https://github.com/Merck/gMCPLite). These two contain only a subset of the original functionality, but they remove the rJava dependency and add plotting functionality based on ggplot2.

However, because development has ceased on the original package, we hope to re-implement the methods with a more general testing framework; with fewer dependencies, in particular shedding the Java dependency; with simpler, more transparent S3 classes; and with improvements to the accuracy of the parametric and Simes test methods.

A portion of Simes testing is also implemented in the lrstat package.

## Citation

```{r citation, results="markup"}
citation("graphicalMCP")
```

## Acknowledgments

We owe a debt of gratitude to the authors of gMCP for their pioneering work in the field, without which this package would not be nearly as extensive as it is.

## Glossary of terms {#glossary-of-terms}

This package seeks to be both accurate and performant, of course. But beyond that, much thought has been put into the readability of the code. Whether being read by a user validating our methods, a developer extending the package to new uses, or a contributor helping enhance the core functionality, we hope that the code contained here can serve as an educational document to grow people's understanding of the graphical approach to multiple comparison procedures.

To that end, there are several entities encountered in the world of graphical MCPs that we define here. Some of these are used only in the internal code of the package, but most are inputs or output in one or more exported functions. If you believe any definitions could be clarified or improved, please contact the package authors or submit an issue to the GitHub repository.

+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| Entity                          | Definition                                                                                                                                                                                                         | Aliases                               | Variable(s)                     | Related                         |
+=================================+====================================================================================================================================================================================================================+=======================================+=================================+=================================+
| **Graph**                       | A set of nodes and edges representing a clinical trial design                                                                                                                                                      |                                       | `graph`                         | Hypotheses, Transitions         |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
|                                 | **Graphs** are so central that two of their core qualities get their own common variable names: Hypothesis names, and number of hypotheses                                                                         |                                       | `hyp_names`, `num_hyps`         |                                 |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Hypotheses**                  | The weighted nodes in a **graph**. Each node represents a null hypothesis corresponding to a treatment endpoint, and its weight the local significance level.                                                      | weights, hypothesis weights           | `hypotheses`                    | Weighting strategy, Transitions |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Transitions**                 | The weighted edges in a **graph**. Each edge defines how to propagate local significance when a source node is deleted.                                                                                            |                                       | `transitions`                   | Hypotheses                      |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Intersection** **hypothesis** | A subset of **hypotheses** from a **graph**. Plural often implies all such subsets.                                                                                                                                | intersection, sub-graph(s), closure   | `intersections`                 | Weighting strategy              |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Weighting strategy**          | The set of all **intersections** and their **weights** according to Algorithm 1 in Bretz et al (2011)                                                                                                              | intersection weights, closure weights | `weighting_strategy`            |                                 |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Adjusted weights**            | The set of **weights**, adjusted according to a testing algorithm:                                                                                                                                                 |                                       | `adjusted_weights`              |                                 |
|                                 |                                                                                                                                                                                                                    |                                       |                                 |                                 |
|                                 | -   Bonferroni: No change                                                                                                                                                                                          |                                       |                                 |                                 |
|                                 | -   Simes: Sum weights for hypotheses with smaller p-values                                                                                                                                                        |                                       |                                 |                                 |
|                                 | -   Parametric: Multiply weights by c-value, based on the joint distribution                                                                                                                                       |                                       |                                 |                                 |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **P-values**                    | The set of p-values from a real or simulated clinical trial                                                                                                                                                        |                                       | `p`                             | Adjusted & ordered p-values     |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Ordered p-values**            | **P-values** sorted from smallest to largest                                                                                                                                                                       |                                       | `ordered_p`                     | (Adjusted) P-values             |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Adjusted p-values**           | **P-values** that have been divided by **adjusted weights**, allowing direct comparison to **alpha** to determine significance                                                                                     |                                       | `adjusted_p`                    | (Ordered) P-values              |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Significance level**          | The threshold chosen for results of a clinical trial to be considered significant                                                                                                                                  |                                       | `alpha`                         | P-values                        |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Test types**                  | A specification of which testing algorithm to use - Bonferroni, Simes, and parametric are supported                                                                                                                | tests                                 | `test_types`                    | Testing strategy                |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Test groups**                 | A partition of nodes in a **graph** specifying which **hypotheses** should be tested together                                                                                                                      | groups                                | `groups`, `test_groups`         | Testing strategy                |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Testing strategy**            | **Test types** and **test groups** combined                                                                                                                                                                        |                                       |                                 |                                 |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Marginal power**              | The power to reject each null **hypothesis** at full **alpha**. Closely related to the non-centrality parameter, which is the mean of each null **hypothesis** in the underlying multivariate normal distribution: |                                       | `marginal_power`                | Correlation matrix              |
|                                 |                                                                                                                                                                                                                    |                                       |                                 |                                 |
|                                 | `ncp = qnorm(1 - alpha) - qnorm(1 - marginal_power)`                                                                                                                                                               |                                       |                                 |                                 |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Correlation matrix**          | Specification of correlations between **hypotheses**. Together with **marginal power**, this specifies the (known or assumed) underlying multivariate normal distribution of the null **hypotheses**.              |                                       | `corr`, `test_corr`, `sim_corr` | Marginal power                  |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Success**                     | A specification of which null **hypotheses** must be rejected to consider a clinical trial a success                                                                                                               |                                       | `sim_success`                   |                                 |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| **Power**                       | Under a given **graph**, **testing strategy**, **significance level**, and underlying distribution, the estimated likelihood that a particular combination of null hypotheses will be rejected                     |                                       | `power_*`                       | Success                         |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| *Delete a **hypothesis***       | Remove a **hypothesis** from a graph, and update the graph according to algorithm 1 of Bretz et al. (2011)                                                                                                         |                                       | N/A                             | Reject a hypothesis             |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
| *Reject a **hypothesis***       | Under a given **graph**, **testing strategy**, and **significance level**, find a hypothesis (clinical endpoint) to be statistically significant, such that the null hypothesis can be rejected                    |                                       |                                 | Delete a hypothesis             |
+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------+---------------------------------+---------------------------------+
