---
title: "validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(graphicalMCP)
library(lrstat)
library(gMCP)
```

# Introduction

graphicalMCP seeks to be the most comprehensive and flexible R package for graphical MCPs, but it is not the first. The packages gMCP and lrstat each cover portions of graphical MCPs. Here we provide a bit of validation against these existing packages.

# Validation

The simple successive graph will be used to demonstrate equivalence between packages.

```{r simple-successive, fig.dim=c(5, 5)}
hypotheses <- c(.5, .5, 0, 0)
transitions <- rbind(
  c(0, .5, .5, 0),
  c(.5, 0, 0, .5),
  c(0, 1, 0, 0),
  c(1, 0, 0, 0)
)

ss_graph <- graph_create(hypotheses, transitions)

plot(ss_graph, vertex.size = 25)
```

## gMCP

gMCP is the pioneer R package for graphical MCPs, and it covers many common use cases. We'll compare it to graphicalMCP for generating the weights of the closure, testing with Bonferroni and parametric testing, and power calculations using Bonferroni testing.

### Generate weights

Weights of the closure are identical, besides a different order and different names.

```{r gen-wgt-gmcp}
graphicalmcp_weights <- graph_generate_weights(ss_graph)
dimnames(graphicalmcp_weights) <- list(NULL, NULL)

gmcp_weights <- generateWeights(ss_graph$transitions, ss_graph$hypotheses)

identical(gmcp_weights, graphicalmcp_weights[15:1, ])
```

### Bonferroni adjusted p-values

```{r test-bonf}
p <- c(.001, .019, .01, .05)

graphicalmcp_test_bonf <- graph_test_shortcut(ss_graph, p)

gmcp_test_bonf <- gMCP(as_graphMCP(ss_graph), p, alpha = .025)

identical(graphicalmcp_test_bonf$outputs$adjusted_p, gmcp_test_bonf@adjPValues)
```

### Parametric adjusted p-values

Parametric adjusted p-values are calculated at different precision between packages.

```{r test-para}
test_corr <- matrix(.5, 4, 4)
diag(test_corr) <- 1

graphicalmcp_test_para <- graph_test_closure(ss_graph, p,
                                             test_types = "p",
                                             test_corr = list(test_corr))

gmcp_test_para <- gMCP(as_graphMCP(ss_graph), p,
                       alpha = .025,
                       correlation = test_corr)

identical(
  round(graphicalmcp_test_para$outputs$adjusted_p, 10),
  round(gmcp_test_para@adjPValues, 10)
)
```

### Bonferroni power

The packages use slightly different simulation methods, which leads to slight differences in power.

```{r power}
marginal_power <- c(.8, .8, .7, .7)

noncentrality_parameter <-
  qnorm(1 - .025, lower.tail = TRUE) -
  qnorm(1 - marginal_power, lower.tail = TRUE)

set.seed(92123)
graphicalmcp_power <- graph_calculate_power(ss_graph,
                                            power_marginal = marginal_power,
                                            sim_corr = test_corr,
                                            sim_n = 2^17)

set.seed(92123)
gmcp_power <- calcPower(ss_graph$hypotheses, .025, ss_graph$transitions,
                        mean = noncentrality_parameter,
                        corr.sim = test_corr,
                        n.sim = 2^17)

print("Power differences")
structure(
  lapply(
    seq_along(gmcp_power),
    function(index) gmcp_power[[index]] - graphicalmcp_power$power[[index]]
  ),
  names = names(graphicalmcp_power$power)[1:4]
)
```

## lrstat

lrstat is not primarily focused on graphical MCPs, but it does include functions to generate weights and run Simes testing, which makes it a good validation complement to gMCP.

### Generate weights

```{r gen-wgt-lrstat}
graphicalmcp_weights <- graph_generate_weights(ss_graph)

lrstat_weights <- fwgtmat(ss_graph$hypotheses, ss_graph$transitions)

identical(lrstat_weights, unname(graphicalmcp_weights[, 5:8]))
```

### Simes adjusted p-values

```{r test-simes}
graphicalmcp_test_simes <- graph_test_closure(ss_graph, p, test_types = "s")

lrstat_test_simes <- fadjpsim(
  fwgtmat(ss_graph$hypotheses, ss_graph$transitions),
  p,
  matrix(1, 1, 4)
)

identical(
  unname(graphicalmcp_test_simes$outputs$adjusted_p),
  lrstat_test_simes
)
```
